---
title: "Indexical Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indexical Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Species Diversity

### Species Richness

### Species Evenness

## Species Diversity Indices

### Simpson’s Index

#### Simpson's Evenness

### Shannon-Weiner Index

#### $H_{max}$

#### Shannon's Evenness

### Berger-Parker Dominance

## {Indexical}

**{Indexical}** is a package designed to help format and analyze biological census data.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
ggplot2::ggplot()
```

```{r setup}
library(indexical)
```

### Functions

#### census()

**census()** tallies observations of each species in your dataset. The **col_name** argument takes the column from your original dataset containing the species name.

census <- function(col_name)

```{r}
c <- census(seaturtles$Species)
print(c)
```

#### div_index() 

**div_index()** calculates several diversity indices (Species Richness, Simpson's Index, Shannon's Index, Berger-Parker Index) and evenness measures (Simpson's Evenness, Shannon's Evenness) from the output dataframe of the **census()** function. The **species_col** argument takes the column containing species names and the **pop_col** takes the column containing population count from the **census()** output dataframe.

div_index <- function(species_col, pop_col)

```{r}
div_ind <- div_index(c$species_names, c$individuals)
print(div_ind)
```

